---
title: "IAPP structures analysis"
output: html_notebook
---

This script can be used to generate all plots in Figures 1, Figure 2 and Supplementary Figure 1. 

Load all packages and functions.

```{r setup}
library(ggplot2)
library(plotly)
library(stats)
library(missMDA)
library(dplyr)
library(ggrepel)
library(viridis)
```

Load per-residue energy contributions. These energies were extracted using the Sequence Detail function in FoldX, applied to Repaired PDB structures along the IAPP fibril maturation trajectory.

```{r}
IAPP_fibrils <- read.csv("Files/Sequence_Detail_Files/Sequence_Detail_IAPP_fibrils.txt", sep="")
list_names <- strsplit(as.character(IAPP_fibrils$Timeorder), "_")
res <- matrix(nrow = length(IAPP_fibrils$Timeorder), ncol = 1)
for (i in 1:length(IAPP_fibrils$Timeorder)){
  res[i,] <- list_names[[i]][1]
}
IAPP_fibrils$Timeorder <- res

#Filter out structures with very low resolutions or high steric clashes
IAPP_resolution <- read.delim("Files/Resolution/IAPP_resolution.txt")
poor_structures <- IAPP_fibrils %>%
  group_by(TO_Chains) %>%
  summarize(avg_value = max(V16))
poor_structures <- merge(poor_structures, IAPP_resolution)
poor_structures$flag <- ifelse(poor_structures$Resolution > (3.5 - 0.125 * poor_structures$avg_value), "Above Line", "Below Line")
poor_structures$clean_labels <- poor_structures$Timeorder
poor_structures$clean_labels <- paste0(poor_structures$Timeorder,"_",poor_structures$Chain)
p <- ggplot(poor_structures, aes(x = avg_value, y = Resolution, color = flag)) +
  geom_point(size = 3) +  # Scatter plot
  # Label only the red points (Above Line)
  geom_text_repel(aes(label = ifelse(flag == "Above Line", clean_labels, NA)), 
                  size = 3, 
                  max.overlaps = Inf) +  

  # Partitioning line
  geom_abline(intercept = 3.5, slope = -0.125, linetype = "dashed", color = "black", size = 1) +  
  # Color customization
  scale_color_manual(values = c("Above Line" = "red", "Below Line" = "black")) +  
  # Labels and theme
  labs(x = "Max. VDW clash value (kcal/mol)", y = "Resolution (Å)", color = "Status") +
  theme_bw(base_size = 14) +  
  theme(
    aspect.ratio = 1,  
    panel.border = element_rect(color = "black", fill = NA, size = 1),  
    # Ensure all text is black
    text = element_text(color = "black"),  
    axis.text.x = element_text(color = "black", angle = 45, hjust = 1, size = 12),  
    axis.text.y = element_text(color = "black", size = 12),  
    axis.title.x = element_text(color = "black", size = 12),  
    axis.title.y = element_text(color = "black", size = 12),  
    plot.subtitle = element_text(color = "black", size = 12),  
    plot.caption = element_text(color = "black", size = 12),  
    legend.text = element_text(color = "black", size = 12),  
    legend.title = element_text(color = "black", size = 12),  
    legend.position = "none"
  )
p
poor_structures$id <- paste0(poor_structures$Timeorder,"_",poor_structures$Chain)
IAPP_resolution$id <- paste0(IAPP_resolution$Timeorder,"_",IAPP_resolution$Chain)
IAPP_resolution <- subset(IAPP_resolution, !id %in% poor_structures$id[which(poor_structures$flag == "Above Line")])
```

Make Ramachandran plots (Supplementary Figure 1).

```{r}
colours <- c("white","darkblue", "lightblue", "yellow","red")
data3 <- IAPP_fibrils
data3$TO_Chains <- NULL
data3$Column <- NULL
data3$CombinedLabel <- with(data3, paste0(Timeorder, " ", Chain))
SSdensity <- ggplot(data3, aes(x = V6, y = V7)) +
  geom_point() + 
  stat_density2d(aes(fill = ..density..^0.20), geom = "tile", contour = FALSE) + 
  scale_fill_gradientn(
    limits = c(0, 0.3),
    colours = colours
  ) +
  scale_x_continuous(expand = c(0, 0), limits = c(-180, 180)) +
  scale_y_continuous(expand = c(0, 0), limits = c(-180, 180)) +
  facet_wrap(~ CombinedLabel, ncol = 7) +
  coord_fixed() +  # <-- this ensures square plots
  ylab("Psi (degrees)") +
  xlab("Phi (degrees)") +
  theme(
    strip.background = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, size = 1)
  )
SSdensity
```

Calculate distribution of dG energies, backbone H-bond energies, solvation energies and side chain burial on core protofibrils of the L-lineage (Figure 1c, Figure 1d, Figure 1e, Figure 1f , Figure 1g and Figure 1h).

```{r}
data3 <- IAPP_fibrils
data3$TO_Chains <- NULL
data3$Column <- NULL
data3$CombinedLabel <- with(data3, paste0(Timeorder, " ", Chain))
data3 <- subset(data3, !Chain %in% c("U-stack", "e-stack"))
data3 <- subset(data3, lineage != "C-lineage")
custom_colors <- c(
  "#05a551",  
  "#f5ee30",  
  "#f7941c",  
  "#a87b2c",  
  "#bf1e2b"   
)
custom_colors <- as.data.frame(custom_colors)
custom_colors$Timeorder <- c("A","B","F","H","I")
data3 <- merge(data3,custom_colors)

#Calculate dg on L-lineage structures (both protofibrils)
p <- ggplot(data3, aes(x = V9, color = CombinedLabel)) +
  geom_density(size = 1.2) +
    scale_color_manual(values = setNames(data3$custom_colors, data3$CombinedLabel)) +  # Use assigned colors 
  geom_vline(xintercept = 0, linetype = "dashed") +
  annotate("text", x = -2.5, y = 0.05, label = "Stabilizing", hjust = 0, size = 4) +
  annotate("text", x = 0.5, y = 0.05, label = "Destabilizing", hjust = 0, size = 4) +
  labs(
    x = expression("dG residue contribution (kcal/mol)"),
    y = "Density"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")
p

#Calculate backbone H-bond on L-lineage structures (both protofibrils)
p <- ggplot(data3, aes(x = V10, color = CombinedLabel)) +
  geom_density(size = 1.2) +
    scale_color_manual(values = setNames(data3$custom_colors, data3$CombinedLabel)) +  # Use assigned colors 
  labs(
    x = expression("dG residue contribution (kcal/mol)"),
    y = "Density"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")
p

#Calculate van der Waals energies on L-lineage structures (both protofibrils)
p <- ggplot(data3, aes(x = V12, color = CombinedLabel)) +
  geom_density(size = 1.2) +
    scale_color_manual(values = setNames(data3$custom_colors, data3$CombinedLabel)) +  # Use assigned colors 
  labs(
    x = expression("dG residue contribution (kcal/mol)"),
    y = "Density"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")
p

#Calculate solvation energies on L-lineage structures (both protofibrils)
p <- ggplot(data3, aes(x = V15, color = CombinedLabel)) +
  geom_density(size = 1.2) +
    scale_color_manual(values = setNames(data3$custom_colors, data3$CombinedLabel)) +  # Use assigned colors 
  labs(
    x = expression("dG residue contribution (kcal/mol)"),
    y = "Density"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")
p

#Calculate backbone burial on L-lineage structures (both protofibrils)
ggplot(data3, aes(x = V34, y = CombinedLabel, fill = CombinedLabel)) +
  geom_density_ridges(scale = 5, color = "black", linewidth = 0.3, alpha = 0.8) +
  scale_fill_manual(values = setNames(data3$custom_colors, data3$CombinedLabel)) +  # Use assigned colors 
  labs(x = "Backbone Burial", y = "Protofibril") +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none")
p

#Backbone H-bond energies for stabilizing and destabilizing residues
data3$is.stable <- "Stabilizing"
data3$is.stable[which(data3$V9 > 0)] <- "Destabilizing"
ggplot(data3, aes(x = is.stable, y = V10, fill = is.stable)) +
  geom_boxplot(notch = TRUE, outlier.color = "black", width = 0.6) +
  scale_fill_manual(values = c("Destabilizing" = "#d9e9a3", "Stabilizing" = "#a6cee3")) +
  labs(y = "Backbone H-bond (kcal/mol)", x = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none"
  ) +
  annotate("text", x = 1.5, y = 0, label = "****", size = 6)
```

Calculate distribution of dG energies, backbone H-bond energies, solvation energies and side chain burial on core protofibrils of the C-lineage (Figure 1c, Figure 1d, Figure 1e, Figure 1f , Figure 1g and Figure 1h).

```{r}
data3 <- IAPP_fibrils
data3$TO_Chains <- NULL
data3$Column <- NULL
data3$CombinedLabel <- with(data3, paste0(Timeorder, " ", Chain))
data3 <- subset(data3, !Chain %in% c("U-stack", "e-stack"))
data3 <- subset(data3, lineage != "L-lineage")
custom_colors <- c(
  "#05a551",  
  "#f9c8d0",  
  "#ec008b",  
  "#662d91",  
  "#2d3092"   
)
custom_colors <- as.data.frame(custom_colors)
custom_colors$Timeorder <- c("A","C","D","E","G")
data3 <- merge(data3,custom_colors)

#Calculate dg on L-lineage structures (both protofibrils)
p <- ggplot(data3, aes(x = V9, color = CombinedLabel)) +
  geom_density(size = 1.2) +
    scale_color_manual(values = setNames(data3$custom_colors, data3$CombinedLabel)) +  # Use assigned colors 
  geom_vline(xintercept = 0, linetype = "dashed") +
  annotate("text", x = -2.5, y = 0.05, label = "Stabilizing", hjust = 0, size = 4) +
  annotate("text", x = 0.5, y = 0.05, label = "Destabilizing", hjust = 0, size = 4) +
  labs(
    x = expression("dG residue contribution (kcal/mol)"),
    y = "Density"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")
p

#Calculate backbone H-bond on L-lineage structures (both protofibrils)
p <- ggplot(data3, aes(x = V10, color = CombinedLabel)) +
  geom_density(size = 1.2) +
    scale_color_manual(values = setNames(data3$custom_colors, data3$CombinedLabel)) +  # Use assigned colors 
  labs(
    x = expression("dG residue contribution (kcal/mol)"),
    y = "Density"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")
p

#Calculate van der Waals energies on L-lineage structures (both protofibrils)
p <- ggplot(data3, aes(x = V12, color = CombinedLabel)) +
  geom_density(size = 1.2) +
    scale_color_manual(values = setNames(data3$custom_colors, data3$CombinedLabel)) +  # Use assigned colors 
  labs(
    x = expression("dG residue contribution (kcal/mol)"),
    y = "Density"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")
p

#Calculate solvation energies on L-lineage structures (both protofibrils)
p <- ggplot(data3, aes(x = V15, color = CombinedLabel)) +
  geom_density(size = 1.2) +
    scale_color_manual(values = setNames(data3$custom_colors, data3$CombinedLabel)) +  # Use assigned colors 
  labs(
    x = expression("dG residue contribution (kcal/mol)"),
    y = "Density"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")
p

#Calculate backbone burial on L-lineage structures (both protofibrils)
ggplot(data3, aes(x = V34, y = CombinedLabel, fill = CombinedLabel)) +
  geom_density_ridges(scale = 5, color = "black", linewidth = 0.3, alpha = 0.8) +
  scale_fill_manual(values = setNames(data3$custom_colors, data3$CombinedLabel)) +  # Use assigned colors 
  labs(x = "Backbone Burial", y = "Protofibril") +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none")
p

#Backbone H-bond energies for stabilizing and destabilizing residues
data3$is.stable <- "Stabilizing"
data3$is.stable[which(data3$V9 > 0)] <- "Destabilizing"
ggplot(data3, aes(x = is.stable, y = V10, fill = is.stable)) +
  geom_boxplot(notch = TRUE, outlier.color = "black", width = 0.6) +
  scale_fill_manual(values = c("Destabilizing" = "#d9e9a3", "Stabilizing" = "#a6cee3")) +
  labs(y = "Backbone H-bond (kcal/mol)", x = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none"
  ) +
  annotate("text", x = 1.5, y = 0, label = "****", size = 6)
```

Per-residue energy profiles (Figure 2A), total stability energies per protofilament (Figure 1B) and individual APR energies per protofilament (Figure 2B)

```{r}
#Start by filtering out the outer chains as well as the two most outer residues
data3 <- IAPP_fibrils
data3_filtered <- subset(data3, !V3 %in% c("A","B","S","T","d","U","n","e"))
data3_filtered$factor <- paste0(data3_filtered$Timeorder,"_",data3_filtered$Chain)
data3_filtered$TO_Chains <- paste0(data3_filtered$TO_Chains,"_",data3_filtered$type)
aggregate_energies <-  aggregate(V9 ~ factor + V4, data=data3_filtered, FUN="mean")
aggregate_energies <- subset(aggregate_energies, !V4 %in% c(10,11))
aggregate_energies2 <- tidyr::spread(aggregate_energies, V4,V9)
rownames(aggregate_energies2) <- aggregate_energies2$factor
aggregate_energies2<-aggregate_energies2[, -1]
a <- Heatmap(as.matrix(aggregate_energies2), 
        name = "DDG (Kcal/mol)", cluster_columns = FALSE, col = circlize::colorRamp2(c(2,0,-2), c("red", "white", "blue")),
        column_names_gp = gpar(fontsize = 12) , row_names_gp = gpar(fontsize = 12),border_gp = gpar(col = "black", lty = 1),
        cluster_rows = FALSE)
a

#Get global energies for each protofilament, as well as individual APR energies. This just displays the energies. Plots were made using Prism
#ALL energies
rowSums(aggregate_energies2, na.rm = TRUE)

#N-ter APR
N_APR <- aggregate_energies2[,1:6]
rowSums(nN_APR, na.rm = TRUE)

#Central APR
M_APR <- aggregate_energies2[,11:16]
rowSums(M_APR, na.rm = TRUE)

#C-ter APR
C_APR <- aggregate_energies2[,19:24]
rowSums(C_APR, na.rm = TRUE)
```

PCA analysis for all IAPP fibrils (Figure 2f and Figure 2g).

```{r}
list_names <- strsplit(as.character(rownames(aggregate_energies2)), "_")
res1 <- matrix(nrow = nrow(aggregate_energies2), ncol = 1)
res2 <- matrix(nrow = nrow(aggregate_energies2), ncol = 1)
for (i in 1:nrow(aggregate_energies2)){
  res1[i,] <- list_names[[i]][1]
  res2[i,] <- list_names[[i]][2]
}
Structure <- res1
Structure <- as.character(res1)
Protofilaments <- res2
group_numbers <- c(1,1,2,2,3,3,4,4,5,6,6,7,7,8,8,9,9,10,10,11,11,12,12,13,13,14,14)
#Data imputation for missing values
imputed_data <- imputePCA(energies_PCA, ncp=2, method = "Regularized")
#New dataframe with imputed data for plotting in Prism
imputed_energies <- as.data.frame(imputed_data$completeObs)
#Run PCA
PCA <- prcomp(imputed_energies, scale. = TRUE)
#Principal components in a dataframe
emb <- as.data.frame(PCA$x)
emb$Structure <- Structure
emb$Protofilaments <- Protofilaments
emb$colnum <- group_numbers
#extract Variances of main PCs -Do this for PC1, PC2 and PC3
PC1_contr<-fviz_contrib(PCA, choice = "var", axes = 1)
PC1_final<-as.data.frame(PC1_contr[1])
PC2_contr<-fviz_contrib(PCA, choice = "var", axes = 2)
PC2_final<-as.data.frame(PC2_contr[1])
PC3_contr<-fviz_contrib(PCA, choice = "var", axes = 3)
PC3_final<-as.data.frame(PC3_contr[1])
PC1_final$PC2 <- PC2_final$data.contrib
PC1_final$PC3 <- PC3_final$data.contrib
PC1_final$data.name <- as.character(PC1_final$data.name)
PC1_final$data.name <- substr(PC1_final$data.name, 2, nchar(PC1_final$data.name))
#Contributions were saved and exported to Prism
# Perform K-Means Clustering
set.seed(123)  # For reproducibility
num_clusters <- 6  # Change this based on the data
kmeans_result <- kmeans(emb[, c("PC1", "PC2", "PC3")], centers = num_clusters)
emb$cluster <- as.factor(kmeans_result$cluster)
emb$colnum <- as.factor(emb$colnum)
# Find max values for each axis
max_PC1 <- max(emb$PC1)
max_PC2 <- max(emb$PC2)
max_PC3 <- max(emb$PC3)
min_PC1 <- min(emb$PC1)
min_PC2 <- min(emb$PC2)
min_PC3 <- min(emb$PC3)
color_palette <- c("#009900", "yellow", "pink", "magenta", "grey", "purple", 
                   "grey", "orange", "darkblue", "grey", "goldenrod4", "grey20", 
                   "darkred", "grey20")

plot <- plot_ly() %>%
  # Main 3D scatter plot
  add_trace(
    data = emb,
    x = ~PC1, 
    y = ~PC2, 
    z = ~PC3,
    #text = ~names,
    color = ~colnum,
    colors = color_palette,
    type = "scatter3d", 
    mode = "markers+text",
    marker = list(size = 8, opacity = 0.8, line = list(color = 'black', width = 1))
  ) %>%
  # XY projection (Z = 0)
  add_trace(
    data = emb,
    x = ~PC1, 
    y = ~PC2, 
    z = -3.7,
    color = ~colnum,
    colors = color_palette,
    type = "scatter3d",
    mode = "markers",
    marker = list(size = 1, opacity = 1, symbol = 'circle')
  ) %>%
  # Add the top axis lines
  # X-axis Top Trace
  add_trace(
    x = c(min_PC1, max_PC1),
    y = c(max_PC2, max_PC2), 
    z = c(max_PC3, max_PC3),
    type = "scatter3d",
    mode = "lines",
    line = list(color = "red", width = 4),
    name = "X-Axis Top"
  ) %>%
  # XZ projection (Y = 0)
  add_trace(
    data = emb,
    x = ~PC1, 
    y = 4, 
    z = ~PC3,
    color = ~colnum,
    colors = color_palette,
    type = "scatter3d",
    mode = "markers",
    marker = list(size = 1, opacity = 1, symbol = 'circle')
  ) %>%
  # YZ projection (X = 0)
  add_trace(
    data = emb,
    x = 4, 
    y = ~PC2, 
    z = ~PC3,
    color = ~colnum,
    colors = color_palette,
    type = "scatter3d",
    mode = "markers",
    marker = list(size = 1, opacity = 1, symbol = 'circle')
  ) %>%
  add_trace(
    x = kmeans_result$centers[, 1],
    y = kmeans_result$centers[, 2],
    z = kmeans_result$centers[, 3],
    type = "scatter3d",
    mode = "markers",
    marker = list(
      opacity = 0.2, 
      size = c(60, 140, 140, 60, 240, 260), 
      color = I(c("#009900", "grey", "orange", "yellow", "pink", "black")),  # Ensure only 6 colors
      symbol = "circle"
    ),
    name = "Cluster Centers"
  ) %>%
  # Customize layout
  layout(
    title = "3D Plot of Secondary Structure Content with 2D Projections",
    scene = list(
      xaxis = list(title = "<b>PC1 (30.5%)</b>", linewidth = 4),
      yaxis = list(title = "<b>PC2 (25.1%)</b>", linewidth = 4),
      zaxis = list(title = "<b>PC3 (11%)</b>", linewidth = 4),
      aspectmode = "manual",
      aspectratio = list(x = 1, y = 1, z = 1),  # Adjust ratio for visibility
      camera = list(eye = list(x = -2.5, y = -1.8, z = 1))
    )
  )
# Show the plot
plot
```
