---
title: "CTE tau structures analysis"
output: html_notebook
---

This script can be used to generate all plots in Figures 4 and Supplementary Figure 4. 

Load all packages and functions.

```{r setup}
library(ggplot2)
library(plotly)
library(stats)
library(missMDA)
library(dplyr)
library(ggrepel)
```

Load per-residue energy contributions. These energies were extracted using the Sequence Detail function in FoldX, applied to Repaired PDB structures along the Chronic Traumatic Encephalopathy (CTE) fibril maturation trajectory.

Only structures passing quality control were retained; models exhibiting excessive steric clashes or derived from very low-resolution experimental data were excluded to ensure reliability of the energy estimates (Supplementary Figure 2b).

```{r}
#Load file and properly order different structures
Sequence_Detail_AD_CTE_fibrils <- read.csv("Files/Sequence_Detail_Files/Sequence_Detail_CTE_fibrils.txt", sep="")
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='MIA-1_Repair.pdb'] <- "A_MIA-1"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='MIA-3_Repair.pdb'] <- "B_MIA-3"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='MIA-4_Repair.pdb'] <- "C_MIA-4"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='MIA-5_Repair.pdb'] <- "D_MIA-5"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='MIA-6_Repair.pdb'] <- "E_MIA-6"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='MIA-7_Repair.pdb'] <- "F_MIA-7"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='MIA-8_Repair.pdb'] <- "G_MIA-8"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='MIA-9_Repair.pdb'] <- "H_MIA-9"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='MIA-10_Repair.pdb'] <- "I_MIA-10"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='MIA-11_Repair.pdb'] <- "J_MIA-11"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='MIA-12_Repair.pdb'] <- "K_MIA-12"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='MIA-13_Repair.pdb'] <- "L_MIA-13"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='MIA-14_Repair.pdb'] <- "M_MIA-14"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='MIA-15_Repair.pdb'] <- "N_MIA-15"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='MIA-18_Repair.pdb'] <- "O_MIA-18"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='LIA-3_Repair.pdb'] <- "P_LIA-3"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='LIA-5_Repair.pdb'] <- "Q_LIA-5"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='LIA-7_Repair.pdb'] <- "R_LIA-7"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='LIA-14_Repair.pdb'] <- "S_LIA-14"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='LIA-4_Repair.pdb'] <- "T_LIA-4"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='LIA-6_Repair.pdb'] <- "U_LIA-6"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='LIA-13_Repair.pdb'] <- "V_LIA-13"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='LIA-17_Repair.pdb'] <- "W_LIA-17"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='CTEI_Repair.pdb'] <- "X_CTEI"
Sequence_Detail_AD_CTE_fibrils$Timeorder2[Sequence_Detail_AD_CTE_fibrils$Structure=='CTEII_Repair.pdb'] <- "Y_CTEII"

#Filter out structures with very low resolutions or high steric clashes
CTE_resolution <- read.delim("Files/Resolution/CTE_resolution.txt")
poor_structures <- Sequence_Detail_AD_CTE_fibrils %>%
  group_by(Timeorder,Chain) %>%
  summarize(avg_value = max(V16))
poor_structures <- merge(poor_structures, CTE_resolution)
poor_structures$flag <- ifelse(poor_structures$Resolution > (3.5 - 0.125 * poor_structures$avg_value), "Above Line", "Below Line")
poor_structures$clean_labels <- gsub("^[A-Z]_", "", poor_structures$Timeorder)
poor_structures$clean_labels <- paste0(poor_structures$clean_labels,"_",poor_structures$Chain)
p <- ggplot(poor_structures, aes(x = avg_value, y = Resolution, color = flag)) +
  geom_point(size = 3) +  # Scatter plot
  # Label only the red points (Above Line)
  geom_text_repel(aes(label = ifelse(flag == "Above Line", clean_labels, NA)), 
                  size = 3, 
                  max.overlaps = Inf) +  
  # Partitioning line
  geom_abline(intercept = 3.5, slope = -0.125, linetype = "dashed", color = "black", size = 1) +  
  # Color customization
  scale_color_manual(values = c("Above Line" = "red", "Below Line" = "black")) +  
  # Labels and theme
  labs(x = "Max. VDW clash value (kcal/mol)", y = "Resolution (Å)", color = "Status") +
  theme_bw(base_size = 14) +  
  theme(
    aspect.ratio = 1,  
    panel.border = element_rect(color = "black", fill = NA, size = 1),  
    # Ensure all text is black
    text = element_text(color = "black"),  
    axis.text.x = element_text(color = "black", angle = 45, hjust = 1, size = 12),  
    axis.text.y = element_text(color = "black", size = 12),  
    axis.title.x = element_text(color = "black", size = 12),  
    axis.title.y = element_text(color = "black", size = 12),  
    plot.subtitle = element_text(color = "black", size = 12),  
    plot.caption = element_text(color = "black", size = 12),  
    legend.text = element_text(color = "black", size = 12),  
    legend.title = element_text(color = "black", size = 12),  
    legend.position = "none"
)
p
poor_structures$id <- paste0(poor_structures$Timeorder,"_",poor_structures$Chain)
Sequence_Detail_AD_CTE_fibrils$id <- paste0(Sequence_Detail_AD_CTE_fibrils$Timeorder,"_",Sequence_Detail_AD_CTE_fibrils$Chain)
Sequence_Detail_AD_CTE_fibrils <- subset(Sequence_Detail_AD_CTE_fibrils, !id %in% poor_structures$id[which(poor_structures$flag == "Above Line")])
```

Make Ramachandran plot (Supplementary Figure 4a).

```{r}
colours <- c("white","darkblue", "lightblue", "yellow","red")
data3 <- Sequence_Detail_AD_CTE_fibrils
data3$TO_Chains <- NULL
data3$Column <- NULL
data3$CombinedLabel <- with(data3, paste0(Timeorder, " ", Chain))
SSdensity <- ggplot(data3, aes(x = V6, y = V7)) +
  geom_point() + 
  stat_density2d(aes(fill = ..density..^0.20), geom = "tile", contour = FALSE) + 
  scale_fill_gradientn(
    limits = c(0, 0.3),
    colours = colours
  ) +
  scale_x_continuous(expand = c(0, 0), limits = c(-180, 180)) +
  scale_y_continuous(expand = c(0, 0), limits = c(-180, 180)) +
  facet_wrap(~ CombinedLabel, ncol = 8) +
  coord_fixed() +  # <-- this ensures square plots
  ylab("Psi (degrees)") +
  xlab("Phi (degrees)") +
  theme(
    strip.background = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, size = 1)
  )
SSdensity
```

Calculate distribution of dG energies, backbone H-bond energies, solvation energies and side chain burial (Figure 4b, Figure 4c, Figure 4d, Figure 4e, Supplementary Figure 4b, Supplementary Figure 4c, Supplementary Figure 4d and Supplementary Figure 4e).

```{r}
library(viridis)
time_order_color <- data.frame(names(table(data3$Timeorder)),viridis(n = 22, option = "A"))
colnames(time_order_color) <- c("Timeorder","Color")
data3 <- merge(data3,time_order_color)

#Calculate dg on subset of structures (all protofibrils)
data33 <- subset(data3, Timeorder %in% c("C_MIA-4","I_MIA-10","N_MIA-15","R_LIA-5","U_LIA-13","Y_CTEII"))
q <- ggplot(data33, aes(x = V9, color = Timeorder)) +
  geom_density(alpha = 0.3, size = 1.1) +  # Add density with transparency
  scale_color_manual(values = setNames(data33$Color, data33$Timeorder)) +  # Use assigned colors
  labs(
    x = "dG residue contribution (kcal/mol)",
    y = "Density"
  ) +
  #xlim(NA, 5) +
  theme_bw(base_size = 14) + 
  theme(
    aspect.ratio = 1,  # Forces a square aspect ratio
    panel.border = element_rect(color = "black", fill = NA, size = 1),  # Adds a box
    plot.title = element_text(face = "bold", size = 20)  # Bold title if needed
  ) +
  theme(
    text = element_text(color = "black"),  # Ensures all text is black
    axis.text.x = element_text(color = "black", angle = 45, hjust = 1, size = 12),  
    axis.text.y = element_text(color = "black", size = 12),  
    axis.title.x = element_text(color = "black", size = 12),  
    axis.title.y = element_text(color = "black", size = 12),  
    plot.title = element_text(color = "black", size = 12),  
    plot.subtitle = element_text(color = "black", size = 12),  
    plot.caption = element_text(color = "black", size = 12),  
    legend.text = element_text(color = "black", size = 12),  
    legend.title = element_text(color = "black", size = 12),  
    legend.position = "none"  # Removes legend
  )
q

#Calculate dg on ALL structures (all protofibrils)
q <- ggplot(data3, aes(x = V9, color = Timeorder)) +
  geom_density(alpha = 0.3, size = 1.1) +  # Add density with transparency
  scale_color_manual(values = setNames(data3$Color, data3$Timeorder)) +  # Use assigned colors
  labs(
    x = "dG residue contribution (kcal/mol)",
    y = "Density"
  ) +
  #xlim(NA, 5) +
  theme_bw(base_size = 14) + 
  theme(
    aspect.ratio = 1,  # Forces a square aspect ratio
    panel.border = element_rect(color = "black", fill = NA, size = 1),  # Adds a box
    plot.title = element_text(face = "bold", size = 20)  # Bold title if needed
  ) +
  theme(
    text = element_text(color = "black"),  # Ensures all text is black
    axis.text.x = element_text(color = "black", angle = 45, hjust = 1, size = 12),  
    axis.text.y = element_text(color = "black", size = 12),  
    axis.title.x = element_text(color = "black", size = 12),  
    axis.title.y = element_text(color = "black", size = 12),  
    plot.title = element_text(color = "black", size = 12),  
    plot.subtitle = element_text(color = "black", size = 12),  
    plot.caption = element_text(color = "black", size = 12),  
    legend.text = element_text(color = "black", size = 12),  
    legend.title = element_text(color = "black", size = 12),  
    legend.position = "none"  # Removes legend
  )

#Quantification of highly destabilizing residues per structure 
data3$dis <- "No"
data3$dis[which(data3$V9 > 1.5)] <- "Yes"
beta_sheet_cor <- as.data.frame(table(data3$Timeorder,data3$dis))
beta_sheet_cor2 <- as.data.frame(table(data3$Timeorder))
beta_sheet_cor <- subset(beta_sheet_cor, Var2 == "Yes" & Freq != 0)
beta_sheet_cor2 <- subset(beta_sheet_cor2, Freq != 0)
colnames(beta_sheet_cor2)[2]<- "Freq2"
beta_sheet_cor <- merge(beta_sheet_cor,beta_sheet_cor2)
beta_sheet_cor$per <- beta_sheet_cor$Freq / beta_sheet_cor$Freq2 * 100
beta_sheet_cor_summary <- beta_sheet_cor %>%
  group_by(Var1) %>%
  summarize(avg_value = mean(per))
beta_sheet_cor_summary$order <- as.numeric(rownames(beta_sheet_cor_summary))
beta_sheet_cor <- merge(beta_sheet_cor,beta_sheet_cor_summary)
# Perform correlation on the averages
cor_result <- cor.test(beta_sheet_cor_summary$order, beta_sheet_cor_summary$avg_value)
# Create the plot
beta_sheet_cor$clean_labels <- gsub("^[A-Z]_", "", beta_sheet_cor$Var1)
colnames(time_order_color)[1] <- "Var1"
beta_sheet_cor <- merge(beta_sheet_cor,time_order_color)
p <- ggplot(beta_sheet_cor, aes(x = clean_labels, y = per, color = clean_labels)) +
  geom_point(alpha = 1, size = 3) +  # Points colored by Var2
  scale_color_manual(values = setNames(beta_sheet_cor$Color, beta_sheet_cor$clean_labels)) +
  geom_smooth(data = beta_sheet_cor_summary, aes(x = order, y = avg_value),
              method = "lm", color = "black", linetype = "dashed", se = FALSE) +  # Dashed correlation line
  labs(
    x = NULL,  # No label for x-axis
    y = "% of highly destabilizing residues (dG > 1.5 kcal/mol)",  # Updated y-axis label
    color = "Protofibril"  # Legend title
  ) +
  scale_x_discrete(limits = unique(beta_sheet_cor$clean_labels)) +  # Set x-axis labels
  annotate("text", x = 1, y = max(beta_sheet_cor$per),  # Adjust position as needed
           label = paste("Correlation:", round(cor_result$estimate, 2), 
                         "\nP-value:", format(cor_result$p.value, digits = 5)),
           hjust = 0, vjust = 1, color = "black", size = 4) +  # Adjust annotation text size (4 = ~12 pt)
  theme_bw(base_size = 14) + 
  theme(
    aspect.ratio = 1,  # Forces a square aspect ratio
    panel.border = element_rect(color = "black", fill = NA, size = 1),  # Adds a box
    plot.title = element_text(face = "bold", size = 20)  # Bold title if needed
  ) +
  theme(
    text = element_text(color = "black"),  # Ensures all text is black
    axis.text.x = element_text(color = "black", angle = 45, hjust = 1, size = 12),  
    axis.text.y = element_text(color = "black", size = 12),  
    axis.title.x = element_text(color = "black", size = 12),  
    axis.title.y = element_text(color = "black", size = 12),  
    plot.title = element_text(color = "black", size = 12),  
    plot.subtitle = element_text(color = "black", size = 12),  
    plot.caption = element_text(color = "black", size = 12),  
    legend.text = element_text(color = "black", size = 12),  
    legend.title = element_text(color = "black", size = 12),  
    legend.position = "none"  # Removes legend
)
p

#Calculate backbone H-bond on subset of structures (all protofibrils)
data33 <- subset(data3, Timeorder %in% c("C_MIA-4","I_MIA-10","N_MIA-15","R_LIA-5","U_LIA-13","Y_CTEII"))
q <- ggplot(data33, aes(x = V10, color = Timeorder)) +
  geom_density(alpha = 0.3, size = 1.1) +  # Add density with transparency
  scale_color_manual(values = setNames(data33$Color, data33$Timeorder)) +  # Use assigned colors
  labs(
    x = "Backbone H-bond (kcal/mol)",
    y = "Density"
  ) +
  #xlim(NA, 5) +
  theme_bw(base_size = 14) + 
  theme(
    aspect.ratio = 1,  # Forces a square aspect ratio
    panel.border = element_rect(color = "black", fill = NA, size = 1),  # Adds a box
    plot.title = element_text(face = "bold", size = 20)  # Bold title if needed
  ) +
  theme(
    text = element_text(color = "black"),  # Ensures all text is black
    axis.text.x = element_text(color = "black", angle = 45, hjust = 1, size = 12),  
    axis.text.y = element_text(color = "black", size = 12),  
    axis.title.x = element_text(color = "black", size = 12),  
    axis.title.y = element_text(color = "black", size = 12),  
    plot.title = element_text(color = "black", size = 12),  
    plot.subtitle = element_text(color = "black", size = 12),  
    plot.caption = element_text(color = "black", size = 12),  
    legend.text = element_text(color = "black", size = 12),  
    legend.title = element_text(color = "black", size = 12),  
    legend.position = "none"  # Removes legend
  )
q

#Calculate backbone H-bond on ALL structures (all protofibrils)
q <- ggplot(data3, aes(x = V10, color = Timeorder)) +
  geom_density(alpha = 0.3, size = 1.1) +  # Add density with transparency
  scale_color_manual(values = setNames(data3$Color, data3$Timeorder)) +  # Use assigned colors
  labs(
    x = "Backbone H-bond (kcal/mol)",
    y = "Density"
  ) +
  #xlim(NA, 5) +
  theme_bw(base_size = 14) + 
  theme(
    aspect.ratio = 1,  # Forces a square aspect ratio
    panel.border = element_rect(color = "black", fill = NA, size = 1),  # Adds a box
    plot.title = element_text(face = "bold", size = 20)  # Bold title if needed
  ) +
  theme(
    text = element_text(color = "black"),  # Ensures all text is black
    axis.text.x = element_text(color = "black", angle = 45, hjust = 1, size = 12),  
    axis.text.y = element_text(color = "black", size = 12),  
    axis.title.x = element_text(color = "black", size = 12),  
    axis.title.y = element_text(color = "black", size = 12),  
    plot.title = element_text(color = "black", size = 12),  
    plot.subtitle = element_text(color = "black", size = 12),  
    plot.caption = element_text(color = "black", size = 12),  
    legend.text = element_text(color = "black", size = 12),  
    legend.title = element_text(color = "black", size = 12),  
    legend.position = "none"  # Removes legend
  )

#Calculate average backbond H-bond energy per structure 
beta_sheet_cor <- data3 %>%
  group_by(Timeorder) %>%
  summarize(avg_value = mean(V10))
# Perform correlation on the averages
beta_sheet_cor$order <- as.numeric(rownames(beta_sheet_cor))
cor_result <- cor.test(beta_sheet_cor$order, beta_sheet_cor$avg_value)
beta_sheet_cor$clean_labels <- gsub("^[A-Z]_", "", beta_sheet_cor$Timeorder)
colnames(time_order_color)[1] <- "Timeorder"
beta_sheet_cor <- merge(beta_sheet_cor,time_order_color)
p <- ggplot(beta_sheet_cor, aes(x = clean_labels, y = avg_value, color = clean_labels)) +
  geom_point(alpha = 1, size = 3) +  # Points colored by Var2
  scale_color_manual(values = setNames(beta_sheet_cor$Color, beta_sheet_cor$clean_labels)) +
  geom_smooth(data = beta_sheet_cor, aes(x = order, y = avg_value),
              method = "lm", color = "black", linetype = "dashed", se = FALSE) +  # Dashed correlation line
  labs(
    x = NULL,  # No label for x-axis
    y = "Average backbone H-bond energy (kcal/mol)",  # Updated y-axis label
    color = "Protofibril"  # Legend title
  ) +
  scale_x_discrete(limits = unique(beta_sheet_cor$clean_labels)) +  # Set x-axis labels
  annotate("text", x = 1, y = max(beta_sheet_cor$avg_value),  # Adjust position as needed
           label = paste("Correlation:", round(cor_result$estimate, 2), 
                         "\nP-value:", format(cor_result$p.value, digits = 5)),
           hjust = 0, vjust = 1, color = "black", size = 4) +  # Adjust annotation text size (4 = ~12 pt)
  theme_bw(base_size = 14) + 
  theme(
    aspect.ratio = 1,  # Forces a square aspect ratio
    panel.border = element_rect(color = "black", fill = NA, size = 1),  # Adds a box
    plot.title = element_text(face = "bold", size = 20)  # Bold title if needed
  ) +
  theme(
    text = element_text(color = "black"),  # Ensures all text is black
    axis.text.x = element_text(color = "black", angle = 45, hjust = 1, size = 12),  
    axis.text.y = element_text(color = "black", size = 12),  
    axis.title.x = element_text(color = "black", size = 12),  
    axis.title.y = element_text(color = "black", size = 12),  
    plot.title = element_text(color = "black", size = 12),  
    plot.subtitle = element_text(color = "black", size = 12),  
    plot.caption = element_text(color = "black", size = 12),  
    legend.text = element_text(color = "black", size = 12),  
    legend.title = element_text(color = "black", size = 12),  
    legend.position = "none"  # Removes legend
)
p

#Calculate Solvation energy on ALL structures
q <- ggplot(data3, aes(x = V15, color = Timeorder)) +
  geom_density(alpha = 0.3, size = 1.1) +  # Add density with transparency
  scale_color_manual(values = setNames(data3$Color, data3$Timeorder)) +  # Use assigned colors
  labs(
    x = "Solvation (kcal/mol)",
    y = "Density"
  ) +
  #xlim(NA, 5) +
  theme_bw(base_size = 14) + 
  theme(
    aspect.ratio = 1,  # Forces a square aspect ratio
    panel.border = element_rect(color = "black", fill = NA, size = 1),  # Adds a box
    plot.title = element_text(face = "bold", size = 20)  # Bold title if needed
  ) +
  theme(
    text = element_text(color = "black"),  # Ensures all text is black
    axis.text.x = element_text(color = "black", angle = 45, hjust = 1, size = 12),  
    axis.text.y = element_text(color = "black", size = 12),  
    axis.title.x = element_text(color = "black", size = 12),  
    axis.title.y = element_text(color = "black", size = 12),  
    plot.title = element_text(color = "black", size = 12),  
    plot.subtitle = element_text(color = "black", size = 12),  
    plot.caption = element_text(color = "black", size = 12),  
    legend.text = element_text(color = "black", size = 12),  
    legend.title = element_text(color = "black", size = 12),  
    legend.position = "none"  # Removes legend
  )
q

#Calculate Side chain burial on ALL structures
q <- ggplot(data3, aes(x = V33, color = Timeorder)) +
  geom_density(alpha = 0.3, size = 1.1) +  # Add density with transparency
  scale_color_manual(values = setNames(data3$Color, data3$Timeorder)) +  # Use assigned colors
  labs(
    x = "Side chain burial",
    y = "Density"
  ) +
  #xlim(NA, 5) +
  theme_bw(base_size = 14) + 
  theme(
    aspect.ratio = 1,  # Forces a square aspect ratio
    panel.border = element_rect(color = "black", fill = NA, size = 1),  # Adds a box
    plot.title = element_text(face = "bold", size = 20)  # Bold title if needed
  ) +
  theme(
    text = element_text(color = "black"),  # Ensures all text is black
    axis.text.x = element_text(color = "black", angle = 45, hjust = 1, size = 12),  
    axis.text.y = element_text(color = "black", size = 12),  
    axis.title.x = element_text(color = "black", size = 12),  
    axis.title.y = element_text(color = "black", size = 12),  
    plot.title = element_text(color = "black", size = 12),  
    plot.subtitle = element_text(color = "black", size = 12),  
    plot.caption = element_text(color = "black", size = 12),  
    legend.text = element_text(color = "black", size = 12),  
    legend.title = element_text(color = "black", size = 12),  
    legend.position = "none"  # Removes legend
  )
q
```

Per-residue energy profiles (Figure 4F)

```{r}
#Start by filtering out the outer chains as well as the two most outer residues
data3_filtered <- subset(data3, !V3 %in% c("A","B","S","T","d","U","n","e"))
data3_filtered$factor <- paste0(data3_filtered$Timeorder,"_",data3_filtered$Chain)
data3_filtered$TO_Chains <- paste0(data3_filtered$TO_Chains,"_",data3_filtered$type)
aggregate_energies <-  aggregate(V9 ~ factor + V4, data=data3_filtered, FUN="mean")
aggregate_energies <- subset(aggregate_energies, !V4 %in% c(304,305,379))
aggregate_energies2 <- tidyr::spread(aggregate_energies, V4,V9)
rownames(aggregate_energies2) <- aggregate_energies2$factor
aggregate_energies2<-aggregate_energies2[, -1]
a <- Heatmap(as.matrix(aggregate_energies2), 
        name = "DDG (Kcal/mol)", cluster_columns = FALSE, col = circlize::colorRamp2(c(3,0,-3), c("red", "white", "blue")),
        column_names_gp = gpar(fontsize = 12) , row_names_gp = gpar(fontsize = 12),border_gp = gpar(col = "black", lty = 1),
        cluster_rows = FALSE)
a
#In the paper, the energies were actually exported to Prism
```

PCA analysis for all CTE fibrils (Figure 4g and Figure 4h).

```{r}
list_names <- strsplit(as.character(rownames(aggregate_energies2)), "_")
res1 <- matrix(nrow = nrow(aggregate_energies2), ncol = 1)
res2 <- matrix(nrow = nrow(aggregate_energies2), ncol = 1)
res3 <- matrix(nrow = nrow(aggregate_energies2), ncol = 1)
for (i in 1:nrow(aggregate_energies2)){
  res1[i,] <- list_names[[i]][1]
  res2[i,] <- list_names[[i]][2]
  res3[i,] <- list_names[[i]][3]
}
Structure <- res1
Structure <- paste0(Structure,"_",res2)
Protofilaments <- res3
group_numbers <- rep(seq_along(rle(Structure)$values), rle(Structure)$lengths)
#Data imputation for missing values
imputed_data <- imputePCA(aggregate_energies2, ncp=2, method = "Regularized")
#New dataframe with imputed data for plotting in Prism
imputed_energies <- as.data.frame(imputed_data$completeObs)
#Run PCA
PCA <- prcomp(imputed_energies, scale. = TRUE)
#Principal components in a dataframe
emb <- as.data.frame(PCA$x)
emb$Structure <- Structure
emb$Protofilaments <- Protofilaments
emb$colnum<- group_numbers
#extract Variances of main PCs -Do this for PC1, PC2 and PC3
PC1_contr<-fviz_contrib(PCA, choice = "var", axes = 1)
PC1_final<-as.data.frame(PC1_contr[1])
PC2_contr<-fviz_contrib(PCA, choice = "var", axes = 2)
PC2_final<-as.data.frame(PC2_contr[1])
PC3_contr<-fviz_contrib(PCA, choice = "var", axes = 3)
PC3_final<-as.data.frame(PC3_contr[1])
# Perform K-Means Clustering
set.seed(123)  # For reproducibility
num_clusters <- 2  # Change this based on the data
kmeans_result <- kmeans(emb[, c("PC1", "PC2", "PC3")], centers = num_clusters)
emb$cluster <- as.factor(kmeans_result$cluster)
emb$colnum <- as.factor(emb$colnum)
# Find max values for each axis
max_PC1 <- max(emb$PC1)
max_PC2 <- max(emb$PC2)
max_PC3 <- max(emb$PC3)
min_PC1 <- min(emb$PC1)
min_PC2 <- min(emb$PC2)
min_PC3 <- min(emb$PC3)
#Get PC axis contributions
fviz_eig(PCA, addlabels = TRUE, ylim = c(0, 90))
fviz_pca_var(PCA, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07")
)
plot <- plot_ly() %>%
  # Main 3D scatter plot
  add_trace(
    data = emb,
    x = ~PC1, 
    y = ~PC2, 
    z = ~PC3,
    #text = ~names,
    hoverinfo = "text",
    type = "scatter3d", 
    mode = "markers+text",
    marker = list(color = ~colnum,
                  colorscale = "Electric",
                  size = 8, 
                  opacity = 0.8, 
                  line = list(color = 'black', width = 1))
  ) %>%
  # XY projection (Z = 0)
  add_trace(
    data = emb,
    x = ~PC1, 
    y = ~PC2, 
    z = min_PC3,
    type = "scatter3d",
    mode = "markers",
    marker = list(color = ~colnum,
                  colorscale = "Electric",
                  size = 1, 
                  opacity = 1, 
                  symbol = 'circle')
  ) %>%
  # XZ projection (Y = 0)
  add_trace(
    data = emb,
    x = ~PC1, 
    y = max_PC2, 
    z = ~PC3,
    type = "scatter3d",
    mode = "markers",
    marker = list(color = ~colnum,
                  colorscale = "Electric",
                  size = 1, 
                  opacity = 1, 
                  symbol = 'circle')
  ) %>%
  # YZ projection (X = 0)
  add_trace(
    data = emb,
    x = max_PC1, 
    y = ~PC2, 
    z = ~PC3,
    type = "scatter3d",
    mode = "markers",
    marker = list(color = ~colnum,
                  colorscale = "Electric",
                  size = 1, 
                  opacity = 1, 
                  symbol = 'circle')
  ) %>%
  # Customize layout
  layout(
    title = "3D Plot of Secondary Structure Content with 2D Projections",
    scene = list(
      xaxis = list(title = "<b>PC1 (24.3%)</b>", linewidth = 4),
      yaxis = list(title = "<b>PC2 (12.1%)</b>", linewidth = 4),
      zaxis = list(title = "<b>PC3 (7.2%)</b>", linewidth = 4, range = list(min_PC3, 5)),
      aspectmode = "manual",
      aspectratio = list(x = 1, y = 1, z = 1),  # Adjust ratio for visibility
      camera = list(
      eye = list(x = 1, y = 1.5, z = 1.5),
      up = list(x = 0, y = 0, z = 1),  # Ensures Z is "up"
      center = list(x = 0, y = 0, z = 0)  # Keeps focus at center
      )
    )
  )
# Show the plot
plot
```


